Dust Clumps In Corotation Around Adolescent M-Dwarfs: Four Years of TESS

### 2023-06-22 paper/ms.tex
update ms

\title{Pre-Main-Sequence M-Dwarfs With Complex Quasiperiodic Variability From Four Years of TESS}

@@

% 250 word max!
\begin{abstract}
  Complex quasiperiodic variables (CQVs) are pre-main-sequence
  M-dwarfs with almost-periodic optical modulation that includes both
  sharp and broad troughs.  The troughs are usually superposed over
  smooth, spot-like modulation.  These objects are typically 10-100
  million years (Myr) old;  they do not have primordial gas-rich
  disks.  Nonetheless, the preferred explanations for their optical
  variability invoke long-lived material in orbit at the Keplerian
  corotation radius, for instance transiting dust clumps or stellar
  prominences.  Here, we report new CQVs discovered in TESS data
  collected between July~2018 and Sep~2022.  From our analysis of
  \nstarssearched\ M-dwarfs with $T$$<$16 and $d$$<$150\,pc, we find
  \ngoods\ gold-standard CQVs.  Most of these discoveries are new, and
  they include the brightest and closest examples of this class of
  object known ($T$$\approx$11; $d$$\approx$20\,pc).  A few objects
  are outliers among outliers; LP 12-502 for instance shows a ``dip
  complex'' with a total duration that is nearly constant over more
  than 1000 cycles, but which exhibits anywhere from four to eight
  local minima per cycle.  LP-502 also demonstrated drastic shape
  changes over less than one cycle, and has up to four distinct
  periods.  More generally, we find that none of the CQVs maintain a
  fixed light curve shape over timescales of more than a few hundred
  cycles -- they are all {\it quasi}periodic.  Although the physical
  explanation behind these objects remains enigmatic, our sample may
  help clarify their origin.  Possible connections to star, disk, and
  exoplanet evolution are emphasized.
\end{abstract}

@@

\subsection{Stellar properties}

@@

\paragraph{Effective temperatures, radii, and masses}
%TODO
{\bf TODO}

@@

The asymmetry is an important characteristic because it can help
diagnose the optical depth of the occulting material as a function of
orbital phase angle.  Sharp leading edges with trailing exponential
egresses for instance have been previously seen for transiting
exocomets and disintegrating rocky bodies
\citep[e.g.][]{2012ApJ...752....1R,2012A&A...545L...5B,2015Natur.526..546V,2019A&A...625L..13Z}.

It is tempting to interpret certain features of the LP 12-502 light
curve in this framework of a disintegrating planetesimal swarm.  For
instance, the complex in cycles 248-298 could be well-fit by a sharp
leading edge that decays over 0.2\,$P$ (3.7\,hours).  The earlier
complex, during cycles 248-264 (pre-state-switch), could be fit by a
similar profile.  There are two degenerate explanations for the small
``wiggles'' in the exponential decay.  In one explanation, the decay
timescale is fast ($\approx$0.03\,$P$), and there are up to five
exocomets, superposed.  In the other, there is only one main
``launching body'', and the material it ejects either has some
sub-structure, or else is passing across (and in synchronicity with) a
star with heavily substructured starspots across its surface.

Perhaps the most glaring issue with the planetesimal swarm idea is
that a large number of the dips show asymmetries in the {\it wrong
direction} relative to the naive expectation of a trailing comet tail.
Dust forward scattering could be one possible way out; so too could
simply invoking non-exponentially decaying dust
distributions as a function of azimuth.

@@

@@

\software{
  %\texttt{arviz} \citep{arviz_2019},
  %\texttt{altaipony} \citep{ilin_flares_2021},
  \texttt{astrobase} \citep{2021zndo...1011188B},
  %\texttt{astroplan} \citep{astroplan2018},
	%\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  %\texttt{astropy} \citep{astropy_2018},
  %\texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  %\texttt{ceres} \citep{brahm_2017_ceres},
  %\texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
  %\texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  %\texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  %dependencies \citep{exoplanet:agol20, exoplanet:kipping13, exoplanet:luger18,
  % 	exoplanet:theano},
	%\texttt{gala} \citep{gala,PriceWhelan_2017_gala_zenodo},
	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
  %\texttt{IPython} \citep{perez_2007},
	%\texttt{isochrones} \citep{morton_2015_isochrones},
	%\texttt{lightkurve} \citep{lightkurve_2018},
  %\texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  %\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  %\texttt{numpy} \citep{walt_numpy_2011}, 
  %\texttt{pandas} \citep{mckinney-proc-scipy-2010},
  %\texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  %\texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  %\texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{2020NatMe..17..261V},
  \texttt{TESS-point}  \citep{2020ascl.soft03001B},
  %\texttt{tesscut} \citep{brasseur_astrocut_2019},
  %\texttt{unpopular} \citep{hattorio_2021_cpm},
  %\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  %\texttt{webplotdigitzer} \citep{rohatgi_2019},
  %\texttt{wotan} \citep{hippke_wotan_2019}.
}
\

\facilities{
 	{\it Astrometry}:
		Gaia \citep{2018A&A...616A...1G,2018A&A...616A...1G}.
 	{\it Imaging}:
    Second Generation Digitized Sky Survey. %,
    %SOAR~(HRCam; \citealt{tokovinin_ten_2018}).
 	%Keck:II~(NIRC2).
 	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
 	%Gemini:North~(`Alopeke; \citealt{scott_nessi_2018,scott_twin_2021}.
 	{\it Spectroscopy}:
	%CTIO1.5$\,$m~(CHIRON; \citealt{tokovinin_chironfiber_2013}),
  %PFS ({\bf CITE}),
	%Tillinghast:1.5m~(TRES).
  %MPG2.2$\,$m~(FEROS; \citealt{kaufer_commissioning_1999}),
	%AAT~(Veloce; \citealt{gilbert_veloce_2018}).
	%AAT~(HERMES; \citealt{lewis_2002_hermers_2df,sheinis_2015_hermes}),
 	Keck:I~(HIRES; \citealt{1994SPIE.2198..362V}).
 	%VLT:Kueyen~(FLAMES; \citealt{pasquini_2002}).
% 	Euler1.2m~(CORALIE),
% 	ESO:3.6m~(HARPS; \citealt{mayor_setting_2003}).
 	{\it Photometry}:
%	  ASTEP:0.40$\,$m (ASTEP400),
% 	CTIO:1.0m (Y4KCam),
% 	Danish 1.54m Telescope,
%	  El Sauce:0.356$\,$m,
% 	Elizabeth 1.0m at SAAO,
% 	Euler1.2m (EulerCam),
%	  Kepler,
% 	Magellan:Baade (MagIC),
% 	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
%   MuSCAT3 \citep{Narita_2020},
% 	NTT,
% 	SOAR (SOI),
 	  TESS \citep{2015JATIS...1a4003R},
% 	TRAPPIST \citep{jehin_trappist_2011},
% 	VLT:Antu (FORS2).
%    ZTF.
}

>>>>

%\title{Pre-Main-Sequence M-Dwarfs With Complex Quasiperiodic Variability From Four Years of TESS}

\title{Dust Clumps In Corotation Around Adolescent M-Dwarfs: Four Years of TESS}

@@

% 250 word max!
\begin{abstract}
  Complex quasiperiodic variables (CQVs) are pre-main-sequence
  M-dwarfs with almost-periodic optical modulation that includes both
  sharp and broad troughs.  The troughs are usually superposed over
  smooth, spot-like modulation.  These objects are typically 10-100
  million years (Myr) old;  they do not have primordial gas-rich
  disks.  Nonetheless, the preferred explanations for their optical
  variability invoke long-lived material in orbit at the Keplerian
  corotation radius, most likely transiting dust clumps.
  Here, we report new CQVs discovered in TESS data
  collected between July~2018 and Sep~2022.  From our analysis of
  \nstarssearched\ M-dwarfs with $T$$<$16 and $d$$<$150\,pc, we find
  \ngoods\ gold-standard CQVs.  Most of these discoveries are new, and
  they include the brightest and closest examples of this class of
  object known ($T$$\approx$11; $d$$\approx$20\,pc).  A few objects
  are outliers among outliers; LP 12-502 for instance shows a ``dip
  complex'' with a total duration that is nearly constant over more
  than 1000 cycles, but which exhibits anywhere from four to eight
  local minima per cycle.  LP-502 also demonstrated drastic shape
  changes over less than one cycle, and (locally) shows up to four distinct
  periods.  More generally, we find that none of the CQVs maintain a
  fixed light curve shape over timescales of more than a few hundred
  cycles -- they are all {\it quasi}periodic.  Our sample will
  help facilitate future modeling and observational efforts aimed at understanding
  these objects, and connecting them to the broader
  contexts of star, disk, and
  exoplanet evolution.
\end{abstract}

@@

\subsection{Stellar properties}
\label{subsec:starprops}

@@

\paragraph{Effective temperatures, radii, and masses}

We determine the stellar effective temperature and radii through SED
fitting; the masses are then determined based on interpolation against
the ``magnetic'' models of \citet{2016A&A...593A..99F}.

For the SED fitting, we use \texttt{astroARIADNE}
\citep{2022MNRAS.513.2719V}, tuned to use only the BT-Settl stellar
atmosphere models \citep{Allard2012} based on the
\citet{2009ARA&A..47..481A} solar abundances.  The broadband
magnitudes we considered included $GG_{\rm BP}G_{\rm RP}$ from Gaia
DR2, $Vgri$ from APASS, $JHK_{\rm S}$ from 2MASS, $riz$ from
Pan-STARRS1, SDSS $riz$, and the WISE 1-2 passbands.  We specifically
omitted UV flux measurements to avoid biasing our fit with any
possible chromospheric UV excess.
\texttt{astroARIADNE} compares the measured broadband flux measurements
against pre-computed model grids, and by default fits for six
parameters:
$\{ T_{\rm eff}, R_\star, A_{\rm V}, \log g, [{\rm Fe/H}], d \}$.
The distance  prior is drawn from \citet{2021AJ....161..147B}.
The surface gravity and metallicity are generally unconstrained.
And finally, given our particular use-case, 
we assumed the following priors for the temperature, stellar size, and
extinction:
\begin{align}
  T_{\rm eff} / {\rm K}    &\sim \mathcal{N}(3000, 1000), \\
  R_\star / {\rm R}_\odot  &\sim \mathcal{T}_{\rm N}(0.5, 0.3, 0.1, 1.0), \\
  A_{\rm V} / {\rm mag}    &\sim \mathcal{U}(0, 0.2),
\end{align}
for $\mathcal{N}$ the Gaussian and $\mathcal{U}$ the uniform
distributions, and $\mathcal{T}(\mu, \sigma, a, b)$ a truncated normal
distribution with mean $\mu$, standard deviation $\sigma$, and lower
and upper bounds $a$ and $b$.
Using \texttt{Dynesty} \citep{2020MNRAS.493.3132S},
we statically sampled the posterior probability assuming the default
Gaussian likelihood, and set a stopping threshold of ${\rm d}\log
\mathcal{Z} < 0.01$, for $\mathcal{Z}$ the evidence.

Given the effective temperatures and stellar radii from the SED fit,
we then estimate the mass by interpolating against the ``magnetic''
models of \citet{2016A&A...593A..99F}.  The need for models that
incorporate some form of magnetic correction is well-documented for
our regime of active, young M-dwarfs
\citep[e.g.][]{2012ApJ...756...47S,2015ApJ...804..146D,2020ApJ...891...29S}.
While one could of course allow a range of magnetic field strengths,
and a corresponding range of fractions for starspot converage
\citep{2020ApJ...891...29S}, we prefer the magnetic
\citet{2016A&A...593A..99F} models based on the simplicity of their
assumption that the surface magnetic field strength will be fixed at
equipartition.
We therefore interpolate our SED-measured temperatures and radii, at
the known cluster-based stellar ages ages, in order to estimate
(model-dependent) stellar masses.
For the two systems without known ages (see below), we omit this mass
estimate, due to the degeneracies in the inference problem.

@@

The asymmetry is an important characteristic because it can help
diagnose the optical depth of the occulting material as a function of
orbital phase angle.  Sharp leading edges with trailing exponential
egresses for instance have been previously seen for transiting
exocomets and disintegrating rocky bodies
\citep[e.g.][]{2012ApJ...752....1R,2012A&A...545L...5B,2015Natur.526..546V,2019A&A...625L..13Z}.
We discuss this possibility below.

@@

\paragraph{A planetesimal swarm at corotation seems unlikely}
It is tempting to interpret certain features of the LP 12-502 light
curve in this framework of a disintegrating planetesimal swarm.  For
instance, the complex in cycles 248-298 could be well-fit by a sharp
leading edge that decays over 0.2\,$P$ (3.7\,hours).  The earlier
complex, during cycles 248-264 (pre-state-switch), could be fit by a
similar profile.  There are two degenerate explanations for the small
``wiggles'' in the exponential decay.  In one explanation, the decay
timescale is fast ($\approx$0.03\,$P$), and there are up to five
exocomets, superposed.  In the other, there is only one main
``launching body'', and the material it ejects either has some
sub-structure, or else is passing across (and in synchronicity with) a
star with heavily substructured starspots across its surface.

However there are two glaring issues with the disintegrating planetesimal swarm idea.
First, a large number of the dips show asymmetries in the {\it wrong
	direction} relative to the naive expectation of a trailing comet tail.
Dust forward scattering might be one possible way out; so too might
invoking non-exponentially decaying dust
distributions as a function of azimuth.
Second, and perhaps more serious, a planetesimal swarm should be expected
to be {\it exactly periodic} over observable timescales, on the presumption that the planetesimals
would be massive enough to not feel any headwind or magnetic field.
This is not what is observed; instead, every dip period
that is observed to be distinct from the fundamental 18.5611\,hr period
is both short-lived, and unique.
This suggests structures much less massive than a 1--100\,km sized
planetesimal,
 more strongly
which could be more strongly influenced by the stellar magnetic field.

@@

\software{
  %\texttt{arviz} \citep{arviz_2019},
  %\texttt{altaipony} \citep{ilin_flares_2021},
  \texttt{astrobase} \citep{2021zndo...1011188B},
  %\texttt{astroplan} \citep{astroplan2018},
	%\texttt{AstroImageJ} \citep{collins_astroimagej_2017},
  %\texttt{astropy} \citep{astropy_2018},
  %\texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  %\texttt{ceres} \citep{brahm_2017_ceres},
  %\texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019},
  %\texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  %\texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  %dependencies \citep{exoplanet:agol20, exoplanet:kipping13, exoplanet:luger18,
  % 	exoplanet:theano},
	%\texttt{gala} \citep{gala,PriceWhelan_2017_gala_zenodo},
	%\texttt{IDL Astronomy User's Library} \citep{landsman_1995},
  %\texttt{IPython} \citep{perez_2007},
	%\texttt{isochrones} \citep{morton_2015_isochrones},
	\texttt{lightkurve} \citep{2018ascl.soft12013L},
  %\texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  %\texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  %\texttt{numpy} \citep{walt_numpy_2011}, 
  %\texttt{pandas} \citep{mckinney-proc-scipy-2010},
  %\texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  %\texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  %\texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{2020NatMe..17..261V},
  \texttt{TESS-point}  \citep{2020ascl.soft03001B},
  %\texttt{tesscut} \citep{brasseur_astrocut_2019},
  %\texttt{unpopular} \citep{hattorio_2021_cpm},
  %\texttt{VESPA} \citep{morton_efficient_2012,vespa_2015},
  %\texttt{webplotdigitzer} \citep{rohatgi_2019},
  %\texttt{wotan} \citep{hippke_wotan_2019}.
}
\

\facilities{
 	{\it Astrometry}:
		Gaia \citep{2018A&A...616A...1G,2018A&A...616A...1G}.
 	{\it Imaging}:
    Second Generation Digitized Sky Survey. %,
    %SOAR~(HRCam; \citealt{tokovinin_ten_2018}).
 	%Keck:II~(NIRC2).
 	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
 	%Gemini:North~(`Alopeke; \citealt{scott_nessi_2018,scott_twin_2021}.
 	{\it Spectroscopy}:
	%CTIO1.5$\,$m~(CHIRON; \citealt{tokovinin_chironfiber_2013}),
  %PFS ({\bf CITE}),
	%Tillinghast:1.5m~(TRES).
  %MPG2.2$\,$m~(FEROS; \citealt{kaufer_commissioning_1999}),
	%AAT~(Veloce; \citealt{gilbert_veloce_2018}).
	%AAT~(HERMES; \citealt{lewis_2002_hermers_2df,sheinis_2015_hermes}),
 	Keck:I~(HIRES; \citealt{1994SPIE.2198..362V}).
 	%VLT:Kueyen~(FLAMES; \citealt{pasquini_2002}).
% 	Euler1.2m~(CORALIE),
% 	ESO:3.6m~(HARPS; \citealt{mayor_setting_2003}).
 	{\it Photometry}:
%	  ASTEP:0.40$\,$m (ASTEP400),
% 	CTIO:1.0m (Y4KCam),
% 	Danish 1.54m Telescope,
%	  El Sauce:0.356$\,$m,
% 	Elizabeth 1.0m at SAAO,
% 	Euler1.2m (EulerCam),
%	  Kepler,
% 	Magellan:Baade (MagIC),
% 	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
%   MuSCAT3 \citep{Narita_2020},
% 	NTT,
% 	SOAR (SOI),
 	  TESS \citep{2015JATIS...1a4003R},
% 	TRAPPIST \citep{jehin_trappist_2011},
% 	VLT:Antu (FORS2).
%    ZTF.
  {\it Broadband photometry}:
    2MASS (CITE).
    APASS (CITE).
		Gaia \citep{2018A&A...616A...1G,2018A&A...616A...1G}.
    Pan-STARRS1 (CITE).
    SDSS (CITE).
    AllWISE (CITE).
}
